---
title: "CFLD NUS Challenge"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Team : Fantastic 4
#Debosmitha Chaudhuri
#Mayank Khanna
#Rahul Gupta
#Swetha Narayan

```{r}
require(foreign)
require(ggplot2)
require(MASS)
require(Hmisc)
require(reshape2)
require(randomForest)
```




```{r}
#Import data
full_data <- read.csv("final_consolidated.csv")
str(full_data)
```




## Declare ordinal Variables and numeric variables


```{r}
full_data$Connectivity.and.Infra <- ordered(full_data$Connectivity.and.Infra,
                     levels = c("poor","Bad","Below Average", "Average","Above Average","Good")
                     )
full_data$Residential.Cost <- ordered(full_data$Residential.Cost,
                                          levels = c("Poor","Bad","Below Average", "Average","Above Average","Good")
)
full_data$Proximity.to.organized.retail <- ordered(full_data$Proximity.to.organized.retail,
                                          levels = c("Poor","Bad","Below Average", "Average","Above Average","Good")
)
full_data$Proximity.to.commercial.dev <- ordered(full_data$Proximity.to.commercial.dev,
                                          levels = c("Poor","Bad","Below Average", "Average","Above Average","Good")
)
full_data$Scope.of.Future.Infra.dev <- ordered(full_data$Scope.of.Future.Infra.dev,
                                          levels = c("Poor","Bad","Below Average", "Average","Above Average","Good")
)
full_data$Scope.of.Future.Emploment.Generation <- ordered(full_data$Scope.of.Future.Emploment.Generation,
                                          levels = c("Poor","Bad","Below Average", "Average","Above Average","Good")
)


full_data$Population <- as.numeric(full_data$Population)
```

#Data Cleaning

```{r}
full_data$State.Code <-NULL
full_data$District.Code <- NULL
full_data$Tehsil.Code <- NULL
full_data$Town.Code.Village.code <-NULL
full_data$Sectors.Localities <- NULL
full_data$Price.range <- NULL
full_data$Ward.No <- NULL
str(full_data)


full_data$from.igia <- ifelse(is.na(full_data$from.igia),full_data$Distance.from.Airport,full_data$from.igia)
full_data$price_range <- ifelse(full_data$Low.Price<=5000,"0-5000",ifelse(full_data$Low.Price<=10000,"5000-1000","Greater than 10000"))

full_data$price_range <- ordered(full_data$price_range,levels = c("0-5000","5000-1000", "Greater than 10000")
)
```

#Exploratory analysis
```{r}
summary(full_data)
mod1 <- lm(data = full_data, Low.Price~Air.quality.Index+Crime.index+Connectivity.and.Infra+ 
             Residential.Cost+Proximity.to.organized.retail+Proximity.to.commercial.dev+Scope.of.Future.Infra.dev+
           Scope.of.Future.Emploment.Generation+Population+Rural.Urban+Total.Household+Total.residence+Total.Residence..other.use+
           No.exclusive.room+One.room+Two.rooms+Three.rooms+Four.rooms+Five.rooms+Six.rooms.and.above+X1.Member.family+
           X2.Member.family+X3.Member.family+X4.Member.family+X5.Member.family+X6.8.Member.Family+X9..Member.family+
           Self.Owned+Rented+Other.Ownership+Tapwater.from.treated.source+Covered.well+Un.covered.well+Handpump+Tubewell.Borehole+
           Spring+River..Canal+Tank..Pond..Lake+Other.Water.Source+Electricity.lighting+Kerosene.lighting+Solar.Energy.lighting+
           Other.oil.lighting+Any.other.lighting+No.lighting+Flush.to.piped.sewer.system+Flush.to.Septic.tank+Flush.to.Other.system+
           Closed.drainage+Open.drainage+No.drainage+state.eodb+reforms.access.to.information+reforms.single.window+
             reforms.availability.of.land+reforms.construction.permit.enablers+reforms.environmental.registration.enablers+
           from.igia)
mod2 <- lm(data = full_data,Low.Price ~ Total.No..of.Houses+Child+Schedule.Caste+Schedule.Tribe+
             Literacy+Total.Workers+Main.Worker+Marginal.Worker)
summary(mod1)
```

#Question 1= Random forest

```{r}
library(randomForest)
rf_data <- full_data[ , colSums(is.na(full_data)) == 0]
dim(rf_data)
summary(rf_data$price_range)
mod2 <- randomForest(data = rf_data, price_range~Air.quality.Index+Crime.index
             +Proximity.to.organized.retail+Scope.of.Future.Infra.dev+
             Scope.of.Future.Emploment.Generation+Population+Rural.Urban+Total.Household+Total.residence+Total.Residence..other.use+
             No.exclusive.room+One.room+Two.rooms+Three.rooms+Four.rooms+Five.rooms+Six.rooms.and.above+X1.Member.family+
             X2.Member.family+X3.Member.family+X4.Member.family+X5.Member.family+X6.8.Member.Family+X9..Member.family+
             Self.Owned+Rented+Other.Ownership+Tapwater.from.treated.source+Covered.well+Un.covered.well+Handpump+Tubewell.Borehole+
             Spring+River..Canal+Tank..Pond..Lake+Other.Water.Source+Electricity.lighting+Kerosene.lighting+Solar.Energy.lighting+
             Other.oil.lighting+Any.other.lighting+No.lighting+Flush.to.piped.sewer.system+Flush.to.Septic.tank+Flush.to.Other.system+
             Closed.drainage+Open.drainage+No.drainage+state.eodb+reforms.access.to.information+reforms.single.window+
             reforms.availability.of.land+reforms.construction.permit.enablers+reforms.environmental.registration.enablers+
             from.igia)
importance(mod2)
varImpPlot(mod2)
```

#Question 1= Elastic net

```{r}
require(glmnet)
data_glm <- full_data[ , colSums(is.na(full_data)) == 0]
x <- model.matrix(data=data_glm, price_range ~ Air.quality.Index+Crime.index
                  +Proximity.to.organized.retail+Scope.of.Future.Infra.dev+
                    Scope.of.Future.Emploment.Generation+Population+Rural.Urban+Total.Household+Total.residence+Total.Residence..other.use+
                    No.exclusive.room+One.room+Two.rooms+Three.rooms+Four.rooms+Five.rooms+Six.rooms.and.above+X1.Member.family+
                    X2.Member.family+X3.Member.family+X4.Member.family+X5.Member.family+X6.8.Member.Family+X9..Member.family+
                    Self.Owned+Rented+Other.Ownership+Tapwater.from.treated.source+Covered.well+Un.covered.well+Handpump+Tubewell.Borehole+
                    Spring+River..Canal+Tank..Pond..Lake+Other.Water.Source+Electricity.lighting+Kerosene.lighting+Solar.Energy.lighting+
                    Other.oil.lighting+Any.other.lighting+No.lighting+Flush.to.piped.sewer.system+Flush.to.Septic.tank+Flush.to.Other.system+
                    Closed.drainage+Open.drainage+No.drainage+state.eodb+reforms.access.to.information+reforms.single.window+
                    reforms.availability.of.land+reforms.construction.permit.enablers+reforms.environmental.registration.enablers+
                    from.igia)
str(data_glm)
dim(data_glm)
head(data_glm)
y <- data_glm$price_range
head(y)
str(data_glm)
set.seed(1)

K <- 10
n <- nrow(x)

fold <- sample(rep(seq(K), length=n))
table(fold)
library(glmnet)
alphas <- seq(0, 1, 0.1)
en2.cv.error <- data.frame(alpha=alphas)
for (i in 1:length(alphas)){
  en2.cv <- cv.glmnet(x, y, alpha=alphas[i], family = "multinomial", type.measure = "class", foldid=fold)
  en2.cv.error[i, "lambda.1se"] <- en2.cv$lambda.1se
  en2.cv.error[i, "error.1se"] <- min(en2.cv$cvm) + en2.cv$cvsd[which.min(en2.cv$cvm)]
}

dim(data_glm)

en2.cv.error

# optimal lambda and alpha
#en.lam <- en.cv.error[which.min(en.cv.error$error.min), "lambda.min"]
#en.alpha <- en.cv.error[which.min(en.cv.error$error.min), "alpha"]

#Choosing the most regularized model
en2.lam <- en2.cv.error[which.min(en2.cv.error$error.1se), "lambda.1se"]
en2.alpha <- en2.cv.error[which.min(en2.cv.error$error.1se), "alpha"]

# plot optimal alpha
plot(en2.cv.error$alpha, en2.cv.error$error.1se, type="l")
abline(v=en2.alpha, lty=2)


# the optimal model
en2.mod <- glmnet(x, y, alpha=en2.alpha, family = "multinomial")
plot(en2.mod, xvar="lambda", label = TRUE)
abline(v=log(en2.lam2), lty=2)
predict(en2.mod, type="coefficient", s=en2.lam, exact=TRUE)

```


Both the above models give same factors. Hence the recommended factors affecting land value are reliable

#Question 2, 3- Data preperation

```{r}
#data preparation
str(full_data)
q2_data <- full_data[which(!is.na(full_data$distance.jewar)),]
dim(q2_data)
q2_data <- q2_data[ , colSums(is.na(q2_data)) == 0]
str(q2_data)


q2_data$jewar_benefit_flag <- ifelse(q2_data$from.igia>q2_data$distance.jewar,1,0)
sum(q2_data$jewar_benefit_flag)

#Creating future airport distance variable
q2_data$future_airport_distance <- ifelse(q2_data$from.igia<q2_data$distance.jewar,q2_data$from.igia,q2_data$distance.jewar)
#splitting them in test and train
N <- nrow(q2_data)
train.index <- sample(1:N, round(4*N/5))
test.index <- - train.index

train <- q2_data[train.index,]
test <- q2_data[test.index,]

dim(train)
#making model matrix for different models
x.train <- model.matrix(data = train, price_range~Air.quality.Index+Crime.index
                  +Proximity.to.organized.retail+Scope.of.Future.Infra.dev+
                    Scope.of.Future.Emploment.Generation+Population+Rural.Urban+Total.Household+Total.residence+Total.Residence..other.use+
                    No.exclusive.room+One.room+Two.rooms+Three.rooms+Four.rooms+Five.rooms+Six.rooms.and.above+X1.Member.family+
                    X2.Member.family+X3.Member.family+X4.Member.family+X5.Member.family+X6.8.Member.Family+X9..Member.family+
                    Self.Owned+Rented+Other.Ownership+Tapwater.from.treated.source+Covered.well+Un.covered.well+Handpump+Tubewell.Borehole+
                    Spring+River..Canal+Tank..Pond..Lake+Other.Water.Source+Electricity.lighting+Kerosene.lighting+Solar.Energy.lighting+
                    Other.oil.lighting+Any.other.lighting+No.lighting+Flush.to.piped.sewer.system+Flush.to.Septic.tank+Flush.to.Other.system+
                    Closed.drainage+Open.drainage+No.drainage+state.eodb+reforms.access.to.information+reforms.single.window+
                    reforms.availability.of.land+reforms.construction.permit.enablers+reforms.environmental.registration.enablers+
                    from.igia)
y.train <- train$price_range


x.test <- model.matrix(data = test, price_range~Air.quality.Index+Crime.index
                       +Proximity.to.organized.retail+Scope.of.Future.Infra.dev+
                         Scope.of.Future.Emploment.Generation+Population+Rural.Urban+Total.Household+Total.residence+Total.Residence..other.use+
                         No.exclusive.room+One.room+Two.rooms+Three.rooms+Four.rooms+Five.rooms+Six.rooms.and.above+X1.Member.family+
                         X2.Member.family+X3.Member.family+X4.Member.family+X5.Member.family+X6.8.Member.Family+X9..Member.family+
                         Self.Owned+Rented+Other.Ownership+Tapwater.from.treated.source+Covered.well+Un.covered.well+Handpump+Tubewell.Borehole+
                         Spring+River..Canal+Tank..Pond..Lake+Other.Water.Source+Electricity.lighting+Kerosene.lighting+Solar.Energy.lighting+
                         Other.oil.lighting+Any.other.lighting+No.lighting+Flush.to.piped.sewer.system+Flush.to.Septic.tank+Flush.to.Other.system+
                         Closed.drainage+Open.drainage+No.drainage+state.eodb+reforms.access.to.information+reforms.single.window+
                         reforms.availability.of.land+reforms.construction.permit.enablers+reforms.environmental.registration.enablers+
                         from.igia)
y.test <- test$price_range

```

#Cross validation to choose best predictors

```{r}
#doing cross validation to find best parameters
K <- 10
n <- nrow(x.train)

fold <- sample(rep(seq(K), length=n))
table(fold)
library(glmnet)
alphas <- seq(0, 1, 0.1)
en2.cv.error <- data.frame(alpha=alphas)
for (i in 1:length(alphas)){
  en2.cv <- cv.glmnet(x.train, y.train, alpha=alphas[i], family = "multinomial", type.measure = "class", foldid=fold)
  en2.cv.error[i, "lambda.1se"] <- en2.cv$lambda.1se
  en2.cv.error[i, "error.1se"] <- min(en2.cv$cvm) + en2.cv$cvsd[which.min(en2.cv$cvm)]
}

en2.lam2 <- en2.cv.error[which.min(en2.cv.error$error.1se), "lambda.1se"]
en2.alpha2 <- en2.cv.error[which.min(en2.cv.error$error.1se), "alpha"]
```

```{r}
#plots
plot(en2.cv.error$alpha, en2.cv.error$error.1se, type="l")
abline(v=en2.alpha2, lty=2)

```


#Optimal glmnet model
```{r}
# the optimal model
en2.mod <- glmnet(x, y, alpha=en2.alpha2, family = "multinomial")
plot(en2.mod, xvar="lambda", label = TRUE)
abline(v=log(en2.lam2), lty=2)
predictions <- predict(en2.mod,s=en2.lam2, newx = x.test, type = "class")
final_df_glm <- data.frame(actual = y.test, predicted = predictions)
```


#Random forest Model
```{r}
#Random forest model
dim(q2_data)
?randomForest
str(q2_data)
mse.rfs <- rep(0, 55)
for(m in 1:55){
  set.seed(12)
  rf <- randomForest(price_range ~ Air.quality.Index+Crime.index
                     +Proximity.to.organized.retail+Scope.of.Future.Infra.dev+
                       Scope.of.Future.Emploment.Generation+Population+Rural.Urban+Total.Household+Total.residence+Total.Residence..other.use+
                       No.exclusive.room+One.room+Two.rooms+Three.rooms+Four.rooms+Five.rooms+Six.rooms.and.above+X1.Member.family+
                       X2.Member.family+X3.Member.family+X4.Member.family+X5.Member.family+X6.8.Member.Family+X9..Member.family+
                       Self.Owned+Rented+Other.Ownership+Tapwater.from.treated.source+Covered.well+Un.covered.well+Handpump+Tubewell.Borehole+
                       Spring+River..Canal+Tank..Pond..Lake+Other.Water.Source+Electricity.lighting+Kerosene.lighting+Solar.Energy.lighting+
                       Other.oil.lighting+Any.other.lighting+No.lighting+Flush.to.piped.sewer.system+Flush.to.Septic.tank+Flush.to.Other.system+
                       Closed.drainage+Open.drainage+No.drainage+state.eodb+reforms.access.to.information+reforms.single.window+
                       reforms.availability.of.land+reforms.construction.permit.enablers+reforms.environmental.registration.enablers+
                       from.igia, data=train , mtry=m)
  mse.rfs[m] <- rf$err.rate[500]
}
summary(rf)
dim(test)
str(test)
plot(1:55, mse.rfs, type="b", xlab="mtry", ylab="OOB Error")
mse.rfs
optm <- which.min(mse.rfs)
rf_final <- randomForest(price_range ~ Air.quality.Index+Crime.index
                               +Proximity.to.organized.retail+Scope.of.Future.Infra.dev+
                                 Scope.of.Future.Emploment.Generation+Population+Rural.Urban+Total.Household+Total.residence+Total.Residence..other.use+
                                 No.exclusive.room+One.room+Two.rooms+Three.rooms+Four.rooms+Five.rooms+Six.rooms.and.above+X1.Member.family+
                                 X2.Member.family+X3.Member.family+X4.Member.family+X5.Member.family+X6.8.Member.Family+X9..Member.family+
                                 Self.Owned+Rented+Other.Ownership+Tapwater.from.treated.source+Covered.well+Un.covered.well+Handpump+Tubewell.Borehole+
                                 Spring+River..Canal+Tank..Pond..Lake+Other.Water.Source+Electricity.lighting+Kerosene.lighting+Solar.Energy.lighting+
                                 Other.oil.lighting+Any.other.lighting+No.lighting+Flush.to.piped.sewer.system+Flush.to.Septic.tank+Flush.to.Other.system+
                                 Closed.drainage+Open.drainage+No.drainage+state.eodb+reforms.access.to.information+reforms.single.window+
                                 reforms.availability.of.land+reforms.construction.permit.enablers+reforms.environmental.registration.enablers+
                                 from.igia, data=train , mtry=optm)
predictions_rf <- predict(rf_final, newdata = test)
table(test$price_range, predictions_rf)
```


#XGBoost model
```{r}
library(xgboost)
dtrain <- xgb.DMatrix(data=x.train, label=y.train)
dtrain
str(dtrain)
head(getinfo(dtrain, 'label'))

xgb.cv1 <- xgb.cv(data = dtrain,  objective="multi:softmax", 
                  nrounds=5000, max_depth=4, eta=0.005, subsample=0.5, colsample_bytree=1,
                  nfold=10, early_stopping_rounds=100, num_class = 4)
xgb.cv1$evaluation_log$test_merror_mean


objective <- "multi:softmax"
cv.fold <- 10
max_depths <- c(1, 2, 4, 6, 8)  # candidates for d
etas <- c(0.01, 0.005, 0.001)  # candidates for lambda
subsamples <- c(0.5, 0.75, 1)
colsamples <- c(0.6, 0.8, 1)
```


#Tuning for best parameters for model
```{r}
set.seed(4321)
tune.out <- data.frame()
for (max_depth in max_depths) {
  for (eta in etas) {
    for (subsample in subsamples) {
      for (colsample in colsamples) {
        # **calculate max n.trees by my secret formula**
        n.max <- round(100 / (eta * sqrt(max_depth)))
        xgb.cv.fit <- xgb.cv(data = dtrain, objective=objective, nfold=cv.fold, early_stopping_rounds=100, verbose=0,
                             nrounds=n.max, max_depth=max_depth, eta=eta, subsample=subsample, colsample_bytree=colsample, num_class = 4)
        n.best <- xgb.cv.fit$best_ntreelimit
        if (objective == "reg:linear") {
          cv.err <- xgb.cv.fit$evaluation_log$test_rmse_mean[n.best]
        } else if (objective == "multi:softmax") {
          cv.err <- xgb.cv.fit$evaluation_log$test_merror_mean[n.best]
        }
        out <- data.frame(max_depth=max_depth, eta=eta, subsample=subsample, colsample=colsample, n.max=n.max, nrounds=n.best, cv.err=cv.err)
        print(out)
        tune.out <- rbind(tune.out, out)
      }
    }
  }
}

tune.out
opt <- which.min(tune.out$cv.err)
nrounds.opt <-tune.out$nrounds[opt]
max_depth.opt <- tune.out$max_depth[opt]
eta.opt <- tune.out$eta[opt]
subsample.opt <- tune.out$subsample[opt]
colsample.opt <- tune.out$colsample[opt]
```

#Optimal XGBoost model
```{r}
xgb2 <- xgboost(data=dtrain, objective="multi:softmax", nround=nrounds.opt, max.depth=max_depth.opt, eta=eta.opt, subsample=subsample.opt, colsample_bytree=colsample.opt, num_class = 4)
yhat.xgb2 <- predict(xgb2, x.test)
yhat.xgb2 <- ifelse(yhat.xgb2==1,"0-5000",ifelse(yhat.xgb2==2,"5000-1000","Greater than 10000"))
table(yhat.xgb2, y.test)
importance_matrix <- xgb.importance(model = xgb2, feature_names = colnames(x.train))
xgb.plot.importance(importance_matrix=importance_matrix)
```

#Recommendations for Question 2, 3

```{r}

test2 <- subset(q2_data,jewar_benefit_flag==1)
test2$from.igia <- test2$future_airport_distance

x.test2 <- model.matrix(data = test2, price_range~Air.quality.Index+Crime.index
                       +Proximity.to.organized.retail+Scope.of.Future.Infra.dev+
                         Scope.of.Future.Emploment.Generation+Population+Rural.Urban+Total.Household+Total.residence+Total.Residence..other.use+
                         No.exclusive.room+One.room+Two.rooms+Three.rooms+Four.rooms+Five.rooms+Six.rooms.and.above+X1.Member.family+
                         X2.Member.family+X3.Member.family+X4.Member.family+X5.Member.family+X6.8.Member.Family+X9..Member.family+
                         Self.Owned+Rented+Other.Ownership+Tapwater.from.treated.source+Covered.well+Un.covered.well+Handpump+Tubewell.Borehole+
                         Spring+River..Canal+Tank..Pond..Lake+Other.Water.Source+Electricity.lighting+Kerosene.lighting+Solar.Energy.lighting+
                         Other.oil.lighting+Any.other.lighting+No.lighting+Flush.to.piped.sewer.system+Flush.to.Septic.tank+Flush.to.Other.system+
                         Closed.drainage+Open.drainage+No.drainage+state.eodb+reforms.access.to.information+reforms.single.window+
                         reforms.availability.of.land+reforms.construction.permit.enablers+reforms.environmental.registration.enablers+
                         from.igia)
y.test2 <- test2$price_range

test2[which(yhat.xgb3==3),"Area.name"]
yhat.xgb3 <- predict(xgb2, x.test2)
y.test2

```

